---
title: "Obtaining user feedback  "
output: html_document
---

```{r, include = FALSE, echo = FALSE}
leanbuild::set_knitr_image_path()
```

# Obtaining user feedback

```{r, fig.alt="This chapter will demonstrate how to: Understand the goals of obtaining user feedback. Construct user feedback methods that inform work on your tool. Create a plan for how to obtain user feedback for your tool.", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1cd434bkLer_CJ04GzpsZwzeEA9gjc5Ho6QimiHPbyEg/edit#slide=id.gd422c5de97_0_60")
```

## The goal of user feedback

How do you know if your code is working? You test it and get feedback!
Similarly, how do you know if your tool is working for your user? Ask them for feedback!

```{r, fig.alt="Tina the Tool Developer and Uri the Tool User both are looking at Tina’s tool that has “error” written all over it with a warning sign and is on fire. Uri the Tool User says Is this supposed to be on fire? Tina the Tool Developer pours a bucket of water on the fire and says Oh! Let me put that out!", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1cd434bkLer_CJ04GzpsZwzeEA9gjc5Ho6QimiHPbyEg/edit#slide=id.gcdcbd8d802_0_173")
```
^[For all cartoons:     
Avataars by https://getavataaars.com/.   
Icons by https://thenounproject.com/ License CC BY-NC-ND 2.0.     
Emojis by OpenMoji License: CC BY-SA 4.0.]

## How to collect informative user feedback

### User contact method

At the most basic level, you need to provide your users a way to alert you if something with your tool is not working appropriately.

It may sound disappointing that a user has found a problem but this is something to be happy about!  

1) This means someone is interested and invested in using your tool!   
2) They've identified a way to improve your tool to make it even better!  

Providing a method of contact to your users doesn't mean you need to give users your personal email. In fact, that is probably not the most practical way to keep user queries organized.

_Example contact method ideas_:

- Have a link to a form for users to submit ([Google forms](https://www.google.com/forms/about/) are free).   
- Direct users to file a [GitHub issue](https://docs.github.com/en/github/managing-your-work-on-github/about-issues).    
- Have a separate email inbox that you have a notification set up for.  
- Have a [Slack channel](https://slack.com/) that you direct users to.  

Whatever method of contact you provide your users, just make sure its something that works for you and your team to respond to.

### Interviews and usability testing

In whatever contact method you settle on, give your users a way to indicate if they are willing to chat with you or do testing to provide even more feedback.

Depending on your time and resources you can do a lot with usability testing. [This excellent article](https://uxstudioteam.com/ux-blog/usability-testing/) by @Csontos2019 takes us through how to conduct usability testing which we will echo the main points of here.

What to use usability testing for:

> - Identifying main issues in the usability of a product
> - Checking if users understand the steps to carry out a task and the navigation
> - Observing how easily and quickly they accomplish tasks
> - Validating the value proposition of an app or website – do your potential customers understand it?

_To summarize the steps laid out by @Csontos2019 for conducting usability research_:

**[Step 1) Plan out your study](https://uxstudioteam.com/ux-blog/usability-testing/#Step_1_Plan_your_study)** - clarify what you want to learn and write a plan and script.

- What main questions about your tool do you have? -- try to narrow down to one or two priority questions that you can design a test for.  
- What tasks would a user participant perform that would best inform you about your questions?  
- Write out a script that you would use to conduct testing.  

_See more advice about writing usability testing scripts here:_    

- [Steve Krug's usability testing script template](http://sensible.com/downloads/test-script-web.pdf) from @Krug2010.  
- [Gitlab's advice about usability testing script](https://about.gitlab.com/handbook/engineering/ux/ux-research-training/writing-usability-testing-script/) from @Gitlab.  
- [Gitlab's usability testing template script](https://docs.google.com/document/d/1_5Qu2JR9QE5LE6cK4eq9yJs-nXv2rlWWifcjacaiWdI/edit).  


**Pilot your test with colleague(s)**  

Before you use your test on a potential user, borrow a colleague or other internal person to run through your plan and script for a rehearsal. Though your colleagues are likely not the same as your typical user, running your testing plan on someone who works with you first will allow you to identify bugs in your testing plan.
Run this pilot test with your colleague as if you are doing the real thing and try to be as prepared as possible!

**[Step 2) Prepare your test](https://uxstudioteam.com/ux-blog/usability-testing/#Step_3_Lead_the_test)**  

- **Pilot your test with colleague(s)**  

[Use this checklist](http://sensible.com/downloads/checklists.pdf) from @Krug2010 to check if you are prepared!

You can use [these prompts](https://sensible.com/downloads/things-a-therapist-would-say.pdf) from @Krug2010 to help you know what to say to help create a non-judgmental atmosphere for testing.

**[Step 3) Find user participants](https://uxstudioteam.com/ux-blog/usability-testing/#Step_2_User_test_participants)**  

Who is interested in available for being your test subjects? You might be able recruit people from setting up a feedback form, but you also may use word of mouth and ask around. You could also try advertising your need for participants through Twitter, LinkedIn or some other social media platform.

Note: [Calendly](https://calendly.com/) is very handy tool for scheduling usability testing sessions!

**Step 4) Run the test!**  

Now that you've rehearsed your test and scheduled a testing session with a user, you are ready to run the tests and collect information.

A very important point from @Csontos2019:

```{r, fig.alt="Assure your participant that they shoulder no blame. Never let the participants helping you with their valuable feedback feel frustrated.", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1cd434bkLer_CJ04GzpsZwzeEA9gjc5Ho6QimiHPbyEg/edit#slide=id.gd2cd8e726d_0_5")
```

**[Step 4) Analyze & Report](https://uxstudioteam.com/ux-blog/usability-testing/#Step_4_Analyze_Report)**  

Sit down with your notes and recordings, look for patterns.

_More reading on usability testing:_

- [How to run moderated usability testing](https://www.hotjar.com/usability-testing/process-examples/) @Hotjar2020.
- [Usability Testing 101](https://www.nngroup.com/articles/usability-testing-101/) @Moran2019.
- See @Krug2010 's list of [downloadable usability resources](https://sensible.com/download-files/).
- See this [case study](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002554) for a user designer conducting user research on a bioinformatics tool [@Pavelin2012].
- For even more reading, you could buy [this book by @Krug2010](https://www.amazon.com/Rocket-Surgery-Made-Easy-Yourself-ebook/dp/B002UXRGNO).

## Exercise 1: Create a plan for user feedback for your tool

Set up a method of contact for your users. We have a mock feedback user form [set up here](https://docs.google.com/forms/d/1erbaH2k8cra0A2GB6W9Da0tqJCT41ZPlCmHXpKRcMLk/edit?usp=sharing). This form shouldn't be used as is, but could be tailored to more specifically ask questions about your tool that you are looking to get information about.

## Exercise 2: Determine a plan for usability testing

Go through [All You Need to Know to Run Successful Usability Testing](https://uxstudioteam.com/ux-blog/usability-testing/) and construct a plan for usability testing following the steps we discussed above from the article [@Csontos2019]. Think about what in your tool you want to test and write a plan for it.
